% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
  10.5pt,
]{article}
\usepackage{xcolor}
\usepackage[margin=0.9in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
  \setmainfont[]{Times New Roman}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{4pt}{1pt}
\textbackslash titleformat*\{\textbackslash section\}\{\normalsize\bfseries\}
\setlength{\parskip}{1pt}
\setlength{\parindent}{0pt}
\usepackage{enumitem}
\setlist{nosep,topsep=0pt}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Conjecture Discovery via Iterative Refinement: An AI Agent for Open Mathematical Problems},
  pdfauthor={Guancheng Pan, Chu Kochen Honors College, Zhejiang University},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Conjecture Discovery via Iterative Refinement: An AI Agent for
Open Mathematical Problems}
\author{Guancheng Pan, Chu Kochen Honors College, Zhejiang University}
\date{April 2026}

\begin{document}
\maketitle

Existing AI mathematics systems perform proof search (AlphaProof),
formalization (Gauss), or algorithm search (FunSearch). The system
described here performs \textbf{mathematical research}: autonomously
refining the right statement to prove, deriving results not in the
literature, proving publication-grade theorems, and closing the loop
with Lean 4 formal verification. Four demos exhibit four capabilities
--- a closed negative result (OP-2), a new explicit constant (Theorem
3), 7-round conjecture refinement (OP-1), and machine-verified
formalization --- spanning two fields: optimization theory and
low-dimensional topology.

\subsection{Demos 1+2 --- A complete upper/lower bound story for
Stochastic Heavy
Ball}\label{demos-12-a-complete-upperlower-bound-story-for-stochastic-heavy-ball}

\textbf{Problem.} Can SHB (fixed momentum \(\beta>0\)) accelerate over
SGD on \(L\)-smooth convex stochastic problems? I.e., can the bias term
beat \(O(LD^2/T)\) --- like Nesterov AC-SA's \(O(LD^2/T^2)\)?

\textbf{OP-2 (lower bound, CLOSED, publication-grade).} \emph{No.} For
every \((\beta,\eta)\) in the Goujaud feasibility region
\(\mathcal F_{K=3}\), \(\exists\) hard instance with SHB error
\(\Omega(LD^2/T+\sigma D/\sqrt T)\) --- strictly worse than AC-SA,
identical to SGD. Three contributions: \textbf{(a)} \emph{rescaled GTD23
cycling} --- lift the Goujaud--Taylor--Dieuleveut 2023 strongly-convex
cycling construction into the non-SC class by flattening one coordinate
while preserving the \(K{=}3\) orbit on the SC subspace; \textbf{(b)}
\emph{Le Cam two-point variance bound} --- produce
\(\Omega(\sigma D/\sqrt T)\) via a coordinate-separable two-point
hypothesis test, correctly handling SHB's adaptive queries; \textbf{(c)}
\emph{positive-measure cycling region} --- characterize the
zero-momentum-init cycling sub-region
\(\mathcal R^* \subset \mathcal F_{K=3}\) (anchor
\(\beta\!\approx\!0.303\)) with
\(\operatorname{Leb}_3(\mathcal R^*) \ge 1.20\!\times\!10^{-4}\),
certified by 216 mpmath-50-digit grid points + closed-form affine
analysis.

\textbf{Theorem 3 (upper bound, NEW).} For every \(\beta<1\),
deterministic last-iterate
\(f(y_T)-f^* \le C(\beta;k)\,LD^2/(T+W(\beta;k)-1)\), with
\(C(\beta) \approx 0.36\,(1-\beta)^{-1.22}\). \textbf{No prior work
gives an explicit closed-form \(C(\beta)\):} Ghadimi--Liang--Zhang 2015
covers only Cesàro averages; Sebbouh--Gower 2021 gives only small-\(o\).
Method: PEP-style SDP + 2-step Lyapunov with \(k\)-step lookahead
anchors; \(\beta^*_{\text{LMI}}(k){=}0.957\!\to\!0.978\!\to\!0.993\),
geometric ratio \(\approx 0.71\) extrapolating to \(\beta^*_\infty=1\).
\textbf{Nine exact rational certificates}
(\(\beta\in\{0,0.1,\ldots,0.5,0.6,0.8\}\), SymPy QQ, PSD symbolic);
CLARABEL numerical up to \(\beta=0.993\); blow-up exponent
\(p\!\in\![1.21,1.25]\) consistent across \(k=0,1,2\).

\textbf{Joint significance.} Matching upper (Theorem 3) and lower (OP-2)
bounds pin the SHB bias term at \(\Theta(LD^2/T)\) --- a definitive
\emph{no} to ``does fixed-momentum SHB accelerate?'', forming a
self-contained SIOPT/Math. Prog. paper.

\subsection{\texorpdfstring{Demo 3 --- OP-1: 7-round conjecture
refinement on the curve complex
\(C_1(S_{g,n})\)}{Demo 3 --- OP-1: 7-round conjecture refinement on the curve complex C\_1(S\_\{g,n\})}}\label{demo-3-op-1-7-round-conjecture-refinement-on-the-curve-complex-c_1s_gn}

\textbf{Open problem (Souto 2007 et al.):} is \(C_1(S_{g,n})\)
contractible for positive-genus surfaces? The agent ran seven autonomous
refinement rounds:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2941}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2353}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2353}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2353}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
\#
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Conjecture
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Evidence
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Key event
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
1 & Universal vertex & 132/138 sample & DISPROVED: 6 \(K_4\)-core
counterexamples \\
2 & Contractible & 64/64 exhaustive & Verified, no proof technique \\
3 & Dismantlable & 378/378 exhaustive & Polat 2002 /
Boulet--Fieux--Jouve 2008 connection \\
4 & Chordal & 259/271 & 12 non-chordal exceptions at \(k\!\ge\!4\) \\
5 & (W4) wheel + (M) metric & 318/318 exhaustive &
\textbf{Chepoi--Osajda 2014 cross-field connection} \\
6 & Hatcher surgery filler & 12/12 non-chordal & level \(k-2\),
\(i(\alpha,\sigma_\alpha)=0\), uniform \\
7 & Chordal-or-cone dichotomy & 383/389 exhaustive &
\(i(\sigma_\alpha,\beta)=i(\alpha,\beta)\) exactly, 49/49 \\
\end{longtable}
}

The Iteration-5 connection to Chepoi--Osajda was discovered
autonomously: a topology-side failure (no universal vertex on
\(S_{1,2}\) at \(k{=}2\)) triggered a graph-theory literature search,
surfacing a theorem on weakly modular graphs that had not previously
been linked to curve-complex contractibility. \textbf{\(S_{1,1}\) is
closed rigorously} via a Stern--Brocot argument inside the same
Bestvina--Brady filtration. The remaining gap is reduced to a
\emph{single} surgery argument --- prove
\(i(\sigma_\alpha,\beta)=i(\alpha,\beta)\) --- empirically exact across
49/49 pairs. The full statement was \emph{not} visible at iteration 1;
it emerged only through structured failure.

\subsection{Demo 4 --- System architecture and Lean
verification}\label{demo-4-system-architecture-and-lean-verification}

\textbf{Discovery loop:}
\texttt{Proposer\ -\textgreater{}\ Rep\ Selector\ -\textgreater{}\ Verifier\ -\textgreater{}\ Explain-Why\ -\textgreater{}\ Tracker\ -\textgreater{}\ Prover\ -\textgreater{}\ Diagnoser\ -\textgreater{}\ Bridge\ -\textgreater{}\ (repeat)}.
Three modules carry the load. \textbf{(i) Representation Registry} (31
validated entries) records distinct formalizations of the same object
with cost-annotated transports; on stall the agent switches
representation rather than restarting. \textbf{(ii) Explain-Why +
Hypothesis Tracker:} on the smallest tractable instance the agent must
articulate \emph{why} a conjecture holds; the working hypothesis is then
stress-tested on larger instances --- turning failures into sharper
successor conjectures. \textbf{(iii) Lean Verification Backend:}
\textbf{13/13 CERTIFIED} on the optlib benchmark; \texttt{OptLib2} is a
self-contained 1371-LOC library with 9 named theorems, all axiom-clean
over \texttt{{[}propext,\ Classical.choice,\ Quot.sound{]}}. On the
hardest item (Nesterov \(O(1/T^2)\)), the agent \textbf{autonomously
synthesized the auxiliary sequence} \texttt{zSeq} with bridge identities
\texttt{xkk1\_eq\_combine} and \texttt{y\_eq\_combine} (closed by
\texttt{match\_scalars\ \textless{};\textgreater{}\ field\_simp\ \textless{};\textgreater{}\ ring})
--- definition synthesis, not just proof search.

\subsection{What distinguishes this from existing
systems}\label{what-distinguishes-this-from-existing-systems}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1379}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1724}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1724}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1724}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1724}}
  >{\centering\arraybackslash}p{(\linewidth - 10\tabcolsep) * \real{0.1724}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
AlphaProof
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Gauss
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
FunSearch
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Liam Price (vibe maths)
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
\textbf{This system}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Task & Proof search & Formalization & Algorithm search & One-shot
prompting & \textbf{Math research} \\
Input & Fixed statement & Informal proof & Objective fn & Open problem &
\textbf{Open problem} \\
New mathematical results & No & No & Yes (kissing \#) & Yes (\#1196) &
\textbf{Yes (OP-2 + Thm 3)} \\
Reproducible end-to-end & Yes & Yes & Yes & \textbf{No} &
\textbf{Yes} \\
Cross-field analogy transfer & No & No & No & No & \textbf{Yes} \\
Learns from failure & No & No & Evolutionary & No & \textbf{Structured
(registry+tracker)} \\
Formal verification & Lean & Lean & No & No & \textbf{Lean (13/13,
axiom-clean)} \\
Matching upper + lower bound & N/A & N/A & N/A & N/A & \textbf{Yes (OP-2
+ Thm 3)} \\
\end{longtable}
}

All code, verification scripts, Lean proofs, and SDP certificates are
reproducible end-to-end. Author: \textbf{Guancheng Pan}, freshman, Chu
Kochen Honors College, Zhejiang University.

\end{document}
